Analysis of Programming Students' Prompts to Identify Patterns of Critical Thinking in Interactions with Generative AI \\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-08-19 paper/paper.tex
updates

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{array}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Analysis of Programming Students' Prompts to Identify Patterns of
Critical Thinking in Interactions with Generative AI \\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

The study by Chan and Hu et al. \cite{chan23} revealed that both undergraduate
and postgraduate students exhibit positive attitudes toward the use of GenAI
in teaching and learning, noting that it enhances the depth of their thinking
and understanding. Additionally, a systematic review by Lo, Hew, and Jong
\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
leading to improved understanding and academic achievement. Cai et al.
\cite{cai23} further observed that ChatGPT enables students to regulate their
learning pace, promoting self-regulation.

@@

To deal with this issue, some researchers have proposed different pedagogical
strategies. Denny et al. \cite{Denny24} introduced the concept of \textit{Prompt
Problems}, where students solve programming exercises by formulating natural
language prompts. Prasad and Aamod \cite{Prasad24} proposed a self-regulated
learning framework using GenAI to solve programming problems. Lauren and Watta
\cite{Lauren23} explored the integration of generative AI with evidence-based
learning strategies in computer science and engineering courses.

@@

Effective strategies are crucial to determinate the best approaches and training
educators to use GenAI tools effectively can be time-consuming and
resource-intensive. Therefore, understanding how students interact with these
tools is essential. By analyzing these interaction patterns, we can assess the
impact of GenAI tools on students' learning outcomes.
This knowledge can then be used to refine the design of GenAI tools, ensuring
they better support pedagogical strategies and enhance overall learning
experiences.

@@

It is important to consider how GenAI tools affect students across different
demographic groups, knowledge areas, and levels of programming experience.
Understanding these variations can provide insights into the impact of GenAI
tools on their learning outcomes, helping to tailor educational approaches that
address the specific needs of diverse student populations.

@@

Critical thinking is a fundamental skill for students to develop, as it enables
them to analyze, evaluate, and synthesize information. This skill is essential
for problem-solving and decision-making, making it a crucial aspect of
educational success, specifically in programming courses. However, the use of AI
tools like ChatGPT has raised concerns about the potential impact on students'
critical thinking abilities \cite{Vargas2023} \cite{cai23} \cite{chan23}.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote.
% If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{array}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Analysis of Programming Students' Prompts to Identify Interactions
Patterns with Generative AI \\
{\footnotesize \textsuperscript{}}
\thanks{The authors would like to thank the Federal Institute of Education,
Science, and Technology of Rio Grande do Sul (IFRS) for the partial financial
support provided for the execution of this research.}
}

@@

The study by Chan and Hu et al. \cite{chan23} revealed that both undergraduate
and postgraduate students exhibit positive attitudes toward the use of GenAI
in teaching and learning, noting that it enhances the depth of their thinking
and understanding. Additionally, a systematic review by Lo, Hew, and Jong
\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
leading to improved understanding and academic achievement. Besides of that,
it was also observed  that ChatGPT enables students to regulate their learning
pace, promoting self-regulation \cite{Baha24} \cite{cai23}.

@@

To deal with this issue, some researchers have proposed different pedagogical
strategies. Denny et al. \cite{Denny24} introduced the concept of \textit{Prompt
Problems}, where students solve programming exercises by formulating natural
language prompts. Prasad and Aamod \cite{Prasad24} proposed a self-regulated
learning framework using GenAI to solve programming problems. Lauren and Watta
\cite{Lauren23} explored the integration of GenAI with evidence-based
learning strategies in computer science and engineering courses.

@@

The study of effective educational strategies in use of GenAI tools in education
context are essential for determining the best training for educators. However,
as these tools gain popularity among students \cite{Dickey24} the urgency to
equip educators with best practices may lag behind their rapid adoption.
Therefore, analyzing students' interaction patterns with GenAI tools is
essential for understanding how these tools are being used and how educators can
leverage them to enhance learning. In addition, these interaction patterns can
create new opportunities to redesign GenAI tools to better support pedagogical
strategies and improve overall learning.

@@

Nevertheless, it is important to consider how GenAI tools affect students across
different demographic groups, knowledge areas, cultural backgrounds \cite{catalan21}
\cite{neo22} and students' prior experience. Thus, this works is limited to a
specific group of students and a specific GenAI tool. The results may not be
generalizable to other populations or tools.

@@

Critical thinking is a fundamental skill for students to develop, as it enables
them to analyze, evaluate, and synthesize information. This skill is essential
for problem-solving and decision-making, making it a crucial aspect of
educational success, specifically in programming courses. However, the use of AI
tools like ChatGPT has raised concerns about the potential impact on students'
critical thinking abilities \cite{Murillo23} \cite{cai23} \cite{chan23}.
